---
title: "Figure 3: Gold Standards"
subtitle: 'E. coli ORBIT 2023'
author: 'Scott H. Saunders'
output:
  html_document:
    theme: cosmo
    highlight: tango
    code_folding: show
    toc: yes
---


# Notes

This figure contains several datatypes from experiments that were used benchmark ORBIT and compare to the lambda Red method. Data figures were made in R notebooks and exported as pdfs. Cosmetic improvements were made in Adobe Illustrator. Note that Figures 3A & 3C were made in Adobe Illustrator.

----

Setup packages and plotting for the notebook:

```{r setup, echo=T, message=FALSE, warning=FALSE}
# Check packages
source("../tools/package_setup.R")
# Load packages
library(tidyverse)
library(cowplot)
library(kableExtra)
# Code display options
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=FALSE, echo = TRUE, message=FALSE, warning=FALSE, fig.align="center", fig.retina = 2)
# Load plotting tools
source("../tools/plotting_tools.R")
#Modify the plot theme
theme_set(theme_notebook())
```

-----

# Fig. 3B - ORBIT vs. Lambda Red efficiency

Read in the ORBIT data.

```{r}

df_gold <- read_csv('../../data/low_throughput_experiments/2022_08_31_gold_stds.csv') %>% #read in csv
  mutate(eff = Kan_count / LB_count) %>% group_by(locus) %>% mutate(avg_eff = mean(eff)) #calculate efficiency and average efficiency for replicates

df_gold %>% kable() %>% kable_styling() %>% scroll_box(height = '250px')

```

Read in the lambda Red data.

```{r}

df_lam <- read_csv('../../data/low_throughput_experiments/2022_09_07_lambda_red_galK_AA_eff.csv') %>% #read in csv
  mutate(eff = Kan_count / LB_count) %>% group_by(locus) %>% mutate(avg_eff = mean(eff)) #calculate efficiency and average efficiency for replicates

df_lam %>% kable() %>% kable_styling() %>% scroll_box(height = '250px')

```

Next we combine the ORBIT and lambda Red data and calculate the ratio of efficiencies:

```{r}
df_lam_gold <- bind_rows(df_gold %>% mutate(exp = 'ORBIT'), df_lam %>% mutate(exp = 'lambda_red')) %>% 
  mutate(locus = factor(locus, levels =c('galK','hisA','metA','leuD','-'))) %>% 
  mutate(exp = factor(exp, levels = c('ORBIT','lambda_red')))

df_lam_gold %>% filter(replicate == 1) %>% 
  select(-LB_count,-Kan_count,-eff) %>% 
  pivot_wider(names_from = exp, values_from = avg_eff) %>% 
  mutate(orbit_lam_ratio = ORBIT / lambda_red)

```

Then we can plot the data:

```{r}

#Calculate negative control values for ORBIT and lambda red
orbit_bg <- (df_lam_gold %>% filter(locus=='-' & exp == 'ORBIT'))$avg_eff[1]
lam_bg <-  (df_lam_gold %>% filter(locus=='-' & exp == 'lambda_red'))$avg_eff[1]

#Plot individual datapoints, mean dots and crossbars and negative control X's
plot_lam_gold <- ggplot(df_lam_gold %>% filter(locus!='-') %>% mutate(bg = ifelse(exp == 'ORBIT',orbit_bg, lam_bg)), aes(x = exp, y = eff , color = locus)) + 
  geom_point(data = . %>% filter(replicate ==1), aes(y = bg), shape = 4, color= 'light gray')+
  stat_summary(fun = 'mean', geom = 'crossbar',width = 0.5, size = 0.25)+
  stat_summary(fun = 'mean', geom = 'point')+
  geom_point(position = position_jitterdodge(dodge.width = 1, jitter.width = 1), shape = 21) + 
  facet_grid(~locus)+
  scale_color_viridis_d()+
  scale_y_log10(labels = scales::label_percent(accuracy = 0.0001), breaks = c(0.000001, 0.0001, 0.01)) + 
  scale_x_discrete(labels = c('ORBIT','λRed'))+
  labs(y = 'Efficiency', x = NULL, fill = 'Strand') + guides(color = 'none') +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
  

plot_lam_gold
```

# Fig. 3D - ORBIT accuracy

These data are accuracy measurements where recombinants were tested for growth on permissive and selective media. 

```{r}
df_gold_acc <- read_csv("../../data/low_throughput_experiments/2022_08_30_gold_std_2_accuracy.csv") %>% 
    mutate(accuracy = 1-(selective_colonies / permissive_colonies)) %>% 
  group_by(locus) %>% 
  mutate(avg_accuracy = mean(accuracy)) %>% 
  mutate(locus = factor(locus, levels = c('galK','hisA','metA','leuD')))

df_gold_acc %>% kable() %>% kable_styling() %>% scroll_box(height = '250px')
```

Now let's plot the data:

```{r}
plot_gold_acc <- ggplot(df_gold_acc, aes(x = locus, y = accuracy , color = locus)) + 
  geom_hline(yintercept = 0, linetype =2, color = 'light gray')+ #negative control accuracy was 0
  stat_summary(fun = 'mean', geom = 'crossbar',width = 0.5, size = 0.25)+
  stat_summary(fun = 'mean', geom = 'point')+
  geom_jitter(width =0.1, height = 0, shape = 21) + 
  scale_y_continuous(labels = scales::label_percent())+
  labs(y = 'Accuracy', x = 'Locus')+
 # facet_grid(~locus)+
  scale_color_viridis_d()

plot_gold_acc

```

# Create Fig. 3

```{r}
theme_set(theme_figure())

fig_3_gold <- plot_grid(plot_lam_gold + guides(color = 'none'), plot_gold_acc + guides(color = 'none'), ncol = 1, rel_heights = c(1,0.66),
                   align = 'hv', axis = 'lr', scale = 0.9,
                   labels = c('B','D'))


fig_3_gold

save_plot("../../figures/r_pdf_figs/main_figs/fig_3_gold_std.pdf", fig_3_gold, base_width = 3.5, base_height = 3.75)
```

```{r}
sessionInfo()
```
