---
title: "Figure 4: Large deletions and genomic integrations"
subtitle: 'E. coli ORBIT 2023'
author: 'Scott H. Saunders'
output:
  html_document:
    theme: cosmo
    highlight: tango
    code_folding: show
    toc: yes
---


# Notes

This figure contains several datatypes from experiments where ORBIT was used for large genomic deletions and integrations. Data figures were made in R notebooks and exported as pdfs. Cosmetic improvements were made in Adobe Illustrator. Note that Figures 4A, 4D & 4F were made in Adobe Illustrator.

----

Setup packages and plotting for the notebook:

```{r setup, echo=T, message=FALSE, warning=FALSE}
# Check packages
source("../tools/package_setup.R")
# Load packages
library(tidyverse)
library(cowplot)
library(kableExtra)
# Code display options
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=FALSE, echo = TRUE, message=FALSE, warning=FALSE, fig.align="center", fig.retina = 2)
# Load plotting tools
source("../tools/plotting_tools.R")
#Modify the plot theme
theme_set(theme_notebook())
```

-----

# Fig. 4B - Deletion size efficiency

The galK deletions were done separately from the other loci, so let's read that data in first.

```{r}

df_gal_sizes <- read_csv('../../data/low_throughput_experiments/2022_11_09_galK_sizes_3_efficiency.csv') %>% #read in csv
  mutate(eff = Kan_count / LB_count) %>% group_by(del_size) %>% mutate(avg_eff = mean(eff)) #calculate efficiency and average efficiency for replicates

#calculate negative control value from galK experiment
gal_sizes_pInt <- (df_gal_sizes %>% filter(del_size == '-'))$avg_eff[1]

df_gal_sizes %>% kable() %>% kable_styling() %>% scroll_box(height = '250px')

```

Now we will read in the data for the other loci (hisA, metA, and leuD).

```{r}

df_aa_sizes <- read_csv('../../data/low_throughput_experiments/2022_09_07_AA_del_sizes_eff.csv') %>% #read in csv
  mutate(eff = Kan_count / LB_count) %>% group_by(locus, del_size) %>% mutate(avg_eff = mean(eff)) #calculate efficiency and average efficiency for replicates

df_aa_sizes %>% kable() %>% kable_styling() %>% scroll_box(height = '250px')

```


```{r}

#combine datasets
df_gal_aa_sizes <- bind_rows(df_gal_sizes %>% filter(del_size!='-') %>% mutate(del_size = as.numeric(del_size)), 
                             df_aa_sizes %>% filter(locus!='galK')
                             ) %>% 
  mutate(locus = factor(locus, levels = c( 'galK','hisA','metA','leuD')))

#Plot with individual datapoints, and mean points connected by lines
plot_gal_aa_sizes <- ggplot(df_gal_aa_sizes, aes(x = del_size, y = eff, color = locus)) + 
  geom_hline(yintercept = gal_sizes_pInt, linetype = 2, color = 'light gray')+
  geom_jitter(shape = 21, width = 0.05, height = 0, alpha = 0.4) + 
  geom_point(data = . %>% filter(replicate==1), aes(y = avg_eff)) + 
  geom_line(data = . %>% filter(replicate==1), aes(y = avg_eff)) + 
  scale_y_log10(labels = scales::label_percent(accuracy = 0.01)) + 
  scale_x_log10(breaks = c(100, 1000, 10000, 50000), labels = c('100 bp', '1 kb', '10 kb', '50 kb'))+
  labs(y = 'Efficiency', x = 'Deletion Size')+
  scale_color_viridis_d()

plot_gal_aa_sizes

```

# Fig. 4C - Deletion size accuracy

Again galK was done separately from the other loci, so let's read in that data first:

```{r}
df_galK_pin <- read_csv("../../data/low_throughput_experiments/2022_11_09_galK_sizes_3_accuracy.csv") %>% 
  mutate(accuracy = 1-(selective_colonies / permissive_colonies)) %>% 
  group_by(locus, deletion_size) %>% 
  mutate(avg_accuracy = mean(accuracy))

df_galK_pin
```

Then we will read in the data for hisA, metA and leuD:

```{r}
df_aa_pin <- read_csv("../../data/low_throughput_experiments/2022_09_07_AA_sizes_pin_plate_count.csv") %>% 
  mutate(accuracy = 1-(selective_colonies / permissive_colonies)) %>% 
  group_by(locus, deletion_size) %>% 
  mutate(avg_accuracy = mean(accuracy)) %>% 
  mutate(locus = factor(locus, levels = c('hisA','metA','leuD')))

df_aa_pin
```

We will combine the data and plot:


```{r}

#Combine data
df_galK_aa_pin <- bind_rows(df_galK_pin %>% filter(condition != 'pInt_only') %>% mutate(deletion_size = as.numeric(deletion_size)), df_aa_pin)

#Get background values
galK_bg <- (df_galK_pin %>% filter(condition == 'pInt_only' & replicate == 1))$avg_accuracy

#Plot with individual datapoints, and mean points connected by lines
plot_del_accuracy <- ggplot(df_galK_aa_pin, aes(x = deletion_size, y = accuracy, color = locus)) + 
  geom_hline(yintercept = 0, linetype = 2, color = 'light gray')+ #background accuracy for AA loci was 0.
    geom_hline(yintercept = galK_bg, linetype = 4, color = 'light gray')+ #background accuracy for galK is shown separately here.
  geom_line(data = . %>% filter(replicate==1), aes(y = avg_accuracy)) + 
  geom_jitter(shape = 21, width = 0.05, height = 0, alpha = 0.4) + 
  geom_point(data = . %>% filter(replicate==1), aes(y = avg_accuracy)) + 
  scale_y_continuous(labels = scales::label_percent(accuracy = 1), limits = c(0,1)) + 
  scale_x_log10(breaks = c(100, 1000, 10000, 50000), labels = c('100 bp', '1 kb', '10 kb', '50 kb'))+
  labs(y = 'Phenotypic accuracy', x = 'Deletion Size')+
  scale_color_viridis_d()

plot_del_accuracy
```

# Fig. 4E - Integrating plasmid size efficiency (violacein operon)

For this experiment different sized fragments from the luxR inducible violacein operon were cloned into an integrating plasmid. Efficiency was measured for each plasmid with the ∆galK TO. Let's read in the data:

```{r}
df_vio <- read_csv("../../data/low_throughput_experiments/2022_10_25_pInt_vio_sizes.csv") %>% 
  mutate(eff = Kan_count / LB_count) %>% 
  group_by(condition, to, plasmid_size) %>% mutate(avg_eff = mean(eff)) #calculate efficiency and average efficiency for replicates

df_vio %>% kable() %>% kable_styling() %>% scroll_box(height = '250px')

```

Now let's plot:

```{r}

plot_vio <- ggplot(df_vio%>% filter(to != 'control'), aes(x = plasmid_size, y = eff)) + 
  geom_jitter(shape = 21, width = 200, height = 0, color = "#440154FF", alpha = 0.4) + 
  geom_jitter(data = df_vio %>% filter(to == 'control'), shape = 4, width = 0, height = 0, color = 'light gray') + #negative control efficiencies for pInt_vioA-C and pInt_luxR_vioA-E are shown here
  geom_point(data = . %>% filter(replicate ==1), aes(y = avg_eff), color = "#440154FF") + 
  geom_line(data = . %>% filter(replicate ==1), aes(y = avg_eff, group = to), color = "#440154FF") + 
  scale_y_log10(limits = c(NA,0.001), labels = scales::label_percent(accuracy = 0.001))+
  scale_x_continuous(breaks = c(2000, 6000, 10000), labels = c('2 kb', '6 kb', '10 kb'))+
  labs(x = 'Integrating plasmid size', y = 'Efficiency')

plot_vio
```

# Create Fig. 4

```{r}
theme_set(theme_figure())

fig_4_sizes <- plot_grid(plot_gal_aa_sizes + guides(color = 'none'), plot_vio, nrow = 1,
                   align = 'hv', axis = 'lr', scale = 1,
                   labels = c('B','E'))

fig_4_sizes <- plot_grid(plot_gal_aa_sizes + guides(color = 'none'),plot_del_accuracy + guides(color = 'none'), 
                         plot_vio,NULL,
                         nrow = 2, ncol = 2,
                   align = 'hv', axis = 'lr', scale = 1,
                   labels = c('B','C','E'))

fig_4_sizes

save_plot("../../figures/r_pdf_figs/main_figs/fig_4_del_int_sizes.pdf", fig_4_sizes, base_width = 5, base_height = 4)
```


```{r}

sessionInfo()

```