---
title: "Figure S11-13:Full mutant library grids"
subtitle: 'E. coli ORBIT 2022'
author: 'Scott H. Saunders'
output:
  html_document:
    theme: cosmo
    highlight: tango
    code_folding: show
    toc: yes
---


# Notes

This figure contains data from many different experiments that were used to optimize the protocol for ORBIT. 

----

Setup packages and plotting for the notebook:

```{r setup, echo=T, message=FALSE, warning=FALSE}
# Check packages
source("../tools/package_setup.R")
# Load packages
library(tidyverse)
library(cowplot)
library(kableExtra)
library(ggrastr)
# Code display options
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=FALSE, echo = TRUE, message=FALSE, warning=FALSE, fig.align="center", fig.retina = 2)
# Load plotting tools
source("../tools/plotting_tools.R")
#Modify the plot theme
theme_set(theme_figure())
```

-----

```{r}

df_read_nearest_down <- read_csv('../../data/seq_data/left_side_TO_libs/df_read_nearest_left.csv')
df_read_nearest_up <- read_csv('../../data/seq_data/right_side_TO_libs/df_read_nearest_right.csv')
df_read_nearest_up_down <- bind_rows(df_read_nearest_up %>% mutate(up_down = 'up'),df_read_nearest_down %>%  mutate(up_down = 'down')) %>% 
  mutate(off_target = ifelse(min_val >500, T, F)) %>% 
  mutate(perfect = ifelse(min_val == 0, T, F))

df_read_nearest_up_down
```

```{r}
df_read_nearest_up_down_org <- df_read_nearest_up_down %>% 
  arrange(lib,gene_id, up_down, perfect, width ) %>% 
  group_by(lib,gene_id, up_down, perfect) %>% 
  mutate(read_num = row_number())

df_read_nearest_up_down_org
```

```{r}
df_TOs <- read_csv("../../data/seq_data/right_side_TO_libs/df_TOs.csv")
df_TOs

```

```{r}
df_genes <- read_delim("../../data/seq_data/All_instances_of_Genes_in_Escherichia_coli_K-12_substr._MG1655.txt") %>% 
  mutate(left_pos = `Left-End-Position`, right_pos = `Right-End-Position`) %>% 
  select(-`Left-End-Position`, -`Right-End-Position`) %>% 
  mutate(fwd = ifelse(Direction =='+', T, F))

df_genes_lib <- left_join(df_TOs %>% select(lib, 'gene_id'='gene'), df_genes, by = c('gene_id'='Genes'))

df_genes_lib
```

```{r}
df_poly <- df_genes_lib %>% 
 #filter(right_pos<20000) %>% 
 pivot_longer(c(left_pos, right_pos), values_to = 'x_pos', names_to = 'left_right') %>% 
 mutate( y_pos_bottom = 10,y_pos_top = 100, y_pos_center=31.623) %>% 
 pivot_longer(c(y_pos_bottom,y_pos_top, y_pos_center), values_to = 'y_pos', names_to = 'top_bottom') %>% 
 mutate(x_pos = ifelse(left_right == 'left_pos' & Direction == '-' & top_bottom !='y_pos_center', x_pos + 50, x_pos)) %>% 
 mutate(x_pos = ifelse(left_right == 'right_pos' & Direction == '+' & top_bottom !='y_pos_center', x_pos - 50, x_pos)) %>% 
 mutate(point_num = case_when(left_right == 'left_pos' & top_bottom == 'y_pos_bottom' ~ 1,
left_right == 'left_pos' & top_bottom == 'y_pos_center' ~ 2,
left_right == 'left_pos' & top_bottom == 'y_pos_top' ~ 3,
left_right == 'right_pos' & top_bottom == 'y_pos_top' ~ 4,
left_right == 'right_pos' & top_bottom == 'y_pos_center' ~ 5,
left_right == 'right_pos' & top_bottom == 'y_pos_bottom' ~ 6)) %>% 
 arrange(gene_id, point_num)

df_poly
```
- gene arrows
- coordinate labels
- y axis counts
- different color or axis for off target

```{r}

 kb_label <- function(x){
 # from s to ns
 lab <- paste(x / 1000)
 }

plot_all <- ggplot()+
  geom_rect(data = df_TOs %>% select(lib, gene_id = gene, left_oligo_pos, right_oligo_pos) %>% filter( lib == 1),
                     aes(xmin = left_oligo_pos, xmax = right_oligo_pos, ymin = 0, ymax = 50), fill = 'light gray')+
  geom_segment(data = df_read_nearest_down %>% filter(lib==1) %>% filter(min_val <500),
               aes(x = start, xend = end, y = width, yend = width ),
               #position = position_jitter(height = 75),
               alpha = 0.01)+
  geom_segment(data = df_read_nearest_up%>% filter(lib==1) %>% filter(min_val <500),
             aes(x = start, xend = end, y = width, yend = width ),
             #position = position_jitter(height = 75),
             alpha = 0.01)+
  scale_x_continuous(labels = scales::label_comma())+
  scale_y_continuous(breaks = c(0,40,80), limits = c(0,100))+
  labs(x = "Genomic position (bp)", y = 'Mapped read lengths (bp)')+
  guides(fill = 'none')+
  facet_wrap(~gene_id, scales = 'free', ncol = 8,strip.position = 'top')

plot_all <- ggplot()+
  geom_polygon(data = df_poly %>% filter(gene_id %in% c('alpA','argR','arsR')), aes(x = x_pos, y = y_pos, group = gene_id),fill = 'light gray')+
  geom_segment(data = df_read_nearest_up_down_org%>% filter(gene_id %in% c('alpA','argR','arsR')) %>% filter(min_val <500),
             aes(x = start, xend = end, y = read_num, yend = read_num, color = perfect), alpha = 0.25)+
  labs(x = "Genomic position (Kb)", y = 'Mapped read lengths (bp)')+
   scale_x_continuous(labels = kb_label)+
  scale_y_log10()+
  guides(fill = 'none')+
  facet_wrap(~gene_id, scales = 'free_x', ncol = 8,strip.position = 'top')

```

# Fig. S11 - IDT oPool

```{r}
plot_lib_0 <- ggplot()+
  geom_polygon(data = df_poly %>% filter(lib == 0), aes(x = x_pos, y = y_pos, group = gene_id),fill = 'light gray')+
    rasterise(geom_segment(data = df_read_nearest_up_down_org%>% filter(lib == 0) %>% filter(min_val <500),
             aes(x = start, xend = end, y = read_num, yend = read_num, color = perfect), alpha = 0.25, linewidth = 0.5), dpi = 2000, dev = "cairo")+
  labs(x = "Genomic position (Kb)", y = 'Read_count')+
   scale_x_continuous(labels = kb_label,breaks = scales::extended_breaks(n=3))+
  scale_y_log10(limits =c(1, NA))+
  guides(fill = 'none')+
  facet_wrap(~gene_id, scales = 'free', ncol = 8,strip.position = 'top') + guides(color = 'none')

plot_lib_0

#save_plot("../../figures/r_pdf_figs/supp_figs/plot_lib_0.pdf",plot_lib_0, base_height = 4, base_width = 8)
```


# Fig. S13 - Lib #1 - short TFs

```{r}
plot_lib_1 <- ggplot()+
  geom_polygon(data = df_poly %>% filter(lib == 1), aes(x = x_pos, y = y_pos, group = gene_id),fill = 'light gray')+
  rasterise(geom_segment(data = df_read_nearest_up_down_org%>% filter(lib == 1) %>% filter(min_val <500),
             aes(x = start, xend = end, y = read_num, yend = read_num, color = perfect), alpha = 0.25, linewidth = 0.5), dpi = 2000, dev = "cairo")+
  labs(x = "Genomic position (Kb)", y = 'Read count')+
   scale_x_continuous(labels = kb_label,breaks = scales::extended_breaks(n=3))+
  scale_y_log10(limits =c(1, NA))+
  guides(fill = 'none')+
  facet_wrap(~gene_id, scales = 'free', ncol = 8,strip.position = 'top') + guides(color = 'none')

plot_lib_1

#save_plot("../../figures/r_pdf_figs/supp_figs/plot_lib_1.pdf",plot_lib_1, base_height = 10, base_width = 8)
```

# Fig. S14 - Lib #2 - long TFs

Note this plot is so large, that the Rmd notebook seems to have trouble displaying it, however the pdf output works fine.

```{r}
plot_lib_2 <- ggplot()+
  geom_polygon(data = df_poly %>% filter(lib == 2), aes(x = x_pos, y = y_pos, group = gene_id),fill = 'light gray')+
  rasterise(
  geom_segment(data = df_read_nearest_up_down_org%>% filter(lib == 2) %>% filter(min_val <500),
             aes(x = start, xend = end, y = read_num, yend = read_num, color = perfect), alpha = 0.25, linewidth = 0.5), 
  dpi = 2000, dev = "cairo")+
  labs(x = "Genomic position (Kb)", y = 'Read_count')+
  scale_x_continuous(labels = kb_label,breaks = scales::extended_breaks(n=3))+
  scale_y_log10(limits =c(1, NA))+
  guides(fill = 'none')+
  facet_wrap(~gene_id, scales = 'free', ncol = 8,strip.position = 'top') + guides(color = 'none')

#plot_lib_2

#save_plot("../../figures/r_pdf_figs/supp_figs/plot_lib_2.pdf",plot_lib_2, base_height = 20, base_width = 8)
```

# Fig. S15 - Lib #3 - sRNAs

```{r}
plot_lib_3 <- ggplot()+
  geom_polygon(data = df_poly %>% filter(lib == 3), aes(x = x_pos, y = y_pos, group = gene_id),fill = 'light gray')+
  rasterise(geom_segment(data = df_read_nearest_up_down_org%>% filter(lib == 3) %>% filter(min_val <500),
             aes(x = start, xend = end, y = read_num, yend = read_num, color = perfect), alpha = 0.25, linewidth = 0.5), dpi = 2000, dev = "cairo")+
  labs(x = "Genomic position (Kb)", y = 'Read_count')+
   scale_x_continuous(labels = kb_label,breaks = scales::extended_breaks(n=3))+
  scale_y_log10(limits =c(1, NA))+
  guides(fill = 'none')+
  facet_wrap(~gene_id, scales = 'free', ncol = 8,strip.position = 'top') + guides(color = 'none')

plot_lib_3

#save_plot("../../figures/r_pdf_figs/supp_figs/plot_lib_3.pdf",plot_lib_3, base_height = 10, base_width = 8)
```


```{r}
sessionInfo()
```